Analysing the Performance of Shallow-Water Models with the Roofline Model

### 2017-09-21 docs/papers/roofline_habakkuk/habakkuk.tex
habakkuk-paper: add/move words to Introduction

@@

@@

@@

A key component of the Roofline model is the Arithmetic Intensity of
the code being executed:
\begin{equation}
AI = \frac{\textrm{No. of floating-point operations}}{\textrm{Bytes fetched from memory}}
\end{equation}
Since HWPC data proved to be unreliable, we calculated this quantity
manually by examining the source code and counting the number of
memory references and arithmetic operations that it contained. In
doing this counting we assume that any references to adjacent array
elements are fetched in a single cache-line and thus only count once.

>>>>

The Roofline Model (RM)~\cite{roofline} has become a very popular way
of examining the performance of a given computational kernel and
contrasting it with the various limits associated with the system upon
which it is executing. However, its simplicity naturally means that
it has limitations.

In this work we examine those limitations in light of two different
shallow-water, finite-difference models. We explore the reasons behind
the limited utility of the RM for these two models and present a
tool, Habakkuk, that uses static code analysis to provide more
realistic performance predictions for Fortran kernels.

@@

The construction of a Roofline Model requires the peak floating-point
operations per second (FLOPS) and memory bandwidth (GB/s) that the
computer hardware is capable of. Although theoretical values for these
quantities may be computed (using the technical specification of
the machine, e.g. the number of floating-point instructions that may
be issued per clock cycle).

information on both the
computer being used and the code being run. For the former,

@@

A key component of the Roofline Model is the Arithmetic Intensity (AI)
of the code being executed:
\begin{equation}
AI = \frac{\textrm{No. of floating-point operations}}{\textrm{Bytes fetched from memory}}
\end{equation}

@@

A key component of the Roofline model is the Arithmetic Intensity of
the code being executed:
\begin{equation}
AI = \frac{\textrm{No. of floating-point operations}}{\textrm{Bytes fetched from memory}}
\end{equation}

Since HWPC data proved to be unreliable, we calculated this quantity
manually by examining the source code and counting the number of
memory references and arithmetic operations that it contained. In
doing this counting we assume that any references to adjacent array
elements are fetched in a single cache-line and thus only count once.
